---
title: "Science Paper Supplementary Material"
author: "Vandvik et al."
date: "27 6 2017"
output: word_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE, dpi = 300)
```

```{r loadStuff, echo = FALSE}
#Load libraries
library("lubridate")
library("tidyverse")
library("ggplot2")
library("gridExtra")

#load data
load("temperaturedata.Rdata")
load("monthlyTemperaturedata.RData")
load("monthlyClimate2.Rdata")

Longterm <- data_frame(T_level = as.factor(c(rep("Boreal", 4), rep("Sub-alpine", 4), rep("Alpine", 4))),
                       P_level = as.factor(c(rep(c(1,2,3,4),3))),
                       P_mean = c(600, 1161, 2044, 2923, 789, 1356, 1848, 3029, 596, 1321, 1925, 2725),
                       T_mean = c(10.3, 10.55, 10.6, 10.78, 9.14, 9.17, 8.77, 8.67, 6.17, 6.45, 5.87, 6.58),
                       logger = "30 year normal")


```


##  Climate data
The climate grid (Figure 1 in main text) was set up on the basis of downscaled climate data (data from met.no). In this supplement we explore weather data patterns measured locally at the SeedClim sites to confirm the climate grid with groud data.

# Temperature
Time series of monthly average temperature for boreal (dark red), sub-alpine (red) and alpine (orange) sites measured at 2m and 0cm above ground and -5cm below ground confirm that there are systematic differences between the sites in temperature (left side), and especially during the four warmest months (right side), over the six years for which we have logged climate data:


```{r, echo=FALSE} 
FigS1A_MonthlyAirTemp <- monthlyTemperaturedata %>% 
  filter(logger %in% c("temp200cm", "tempabove", "tempsoil")) %>%
  filter(!is.na(T_level)) %>% 
  mutate(T_level = factor(T_level, levels = c("Boreal","Sub-alpine","Alpine"))) %>% 
  mutate(logger = plyr::mapvalues(logger, c("temp200cm", "tempabove", "tempsoil"), c("Air temperature 2m", "Ground temperature 0cm", "Soil temperature -5cm"))) %>% 
  group_by(T_level, date, logger) %>% 
  summarise(mean = mean(value, na.rm = TRUE)) %>% 
  ggplot(aes(x = date, y = mean, color = T_level)) +
  geom_line(size = 0.9) +
  labs(x = "", y = "Average monthly temperature in °C") +
  scale_color_manual(name = "Temperature level", values = c("darkred", "red", "orange"), guide = FALSE) +
  facet_wrap(~ logger, nrow = 3) +
  theme_minimal(base_size = 15)

# Boxplots
FigS1B_MonthlyAirTemp <- monthlyTemperaturedata %>% 
  filter(logger %in% c("temp200cm", "tempabove", "tempsoil")) %>%
  filter(month(date) %in% 6:9) %>% 
  filter(!is.na(T_level)) %>% 
  mutate(T_level = factor(T_level, levels = c("Boreal","Sub-alpine","Alpine"))) %>% 
  mutate(logger = plyr::mapvalues(logger, c("temp200cm", "tempabove", "tempsoil"), c("Air temperature 2m", "Ground temperature 0cm", "Soil temperature -5cm"))) %>% 
  ggplot(aes(x = T_level, y = value, fill = T_level)) +
  geom_boxplot() +
  labs(x = "", y = "Average summer temperature in °C") +
  scale_fill_manual(name = "Temperature level", values = c("darkred", "red", "orange")) +
  facet_wrap(~ logger, nrow = 3) +
  theme_minimal(base_size = 15) +
  theme(legend.position="bottom")


#extract legend
#https://github.com/hadley/ggplot2/wiki/Share-a-legend-between-two-ggplot2-graphs
g_legend <- function(a.gplot){
  tmp <- ggplot_gtable(ggplot_build(a.gplot))
  leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
  legend <- tmp$grobs[[leg]]
  return(legend)}

mylegend <- g_legend(FigS1B_MonthlyAirTemp)

p3 <- grid.arrange(arrangeGrob(FigS1A_MonthlyAirTemp,
                               FigS1B_MonthlyAirTemp + theme(legend.position="none"),
                               nrow=1), mylegend, nrow=2,heights=c(10, 1))

```

The differences are less clear during winter for ground and soil measurements, which is partly due to snow cover. The temperature at the ground and soil level is close to zero and can fall below zero when snow cover is lacking.


# Precipitation
The precipitation data are much noisier than the temperature data (more missing values and outliers), due to logger failure and also because precipitation falls as snow for a large part of the year. Time series of sum of monthly precipitation for the 5 years of the experiment (www.met.no), averaged across sites for each precipitation level according to increasing precipitation levels (1 = dotted, 2 = dashed, 3 = long-dashed and 4 = solid line from light to dark blue) confirm systematic differences between the sites in precipitation (left side). The same trend across precipitaiton level can be seen for the sum of annual precipitation across the 5 years of the experiment (right side): 

```{r, echo=FALSE} 
FigS2A_MonthlyGriddedPrecip <- monthlyClimate2 %>% 
  filter(logger == "Precipitation") %>% 
  group_by(P_level, date) %>% 
  summarise(mean = mean(value, na.rm = TRUE)) %>% 
  ggplot(aes(x = date, y = mean, color = P_level, linetype = P_level)) +
  geom_line(size = 0.9) +
  labs(x = "", y = "Average monthly precipitation in mm") +
  scale_color_manual(name = "Precipitation level", values = c("#DFEBF7", "#9DC3E6", "#2E75B6", "#213964"), guide = FALSE) +
  scale_linetype_manual(name = "Precipitation level", values = c(3, 2, 5, 1), guide = FALSE) + #"dotted", "dashed", "long-dashed", "solid"
  theme_minimal(base_size = 15)


# Boxplots
FigS2B_MonthlyGriddedPrecip <- monthlyClimate2 %>% 
  filter(logger == "Temperature" & month(date) %in% 6:9 | logger == "Precipitation") %>% 
  filter(year(date) != 2016) %>% 
  group_by(logger, T_level, P_level, year(date)) %>% 
  summarise(n = n(), mean = mean(value, na.rm = TRUE), sum = sum(value, na.rm = TRUE)) %>% 
  mutate(mean = ifelse(logger == "Precipitation", sum, mean)) %>% 
  #group_by(logger, P_level) %>% 
  #summarise(N = n(), Mean = mean(mean), SE = sd(mean, na.rm = TRUE) / sqrt(N)) %>% 
  filter(logger == "Precipitation") %>% 
  ggplot(aes(x = P_level, y = sum, fill = P_level)) +
  geom_boxplot() +
  labs(x = "", y = "Average annual precipitation in mm") +
  scale_fill_manual(name = "Precipitation level", values = c("#DFEBF7", "#9DC3E6", "#2E75B6", "#213964")) +
  theme_minimal(base_size = 15) +
  theme(legend.position="bottom")

mylegend <- g_legend(FigS2B_MonthlyGriddedPrecip)

p4 <- grid.arrange(arrangeGrob(FigS2A_MonthlyGriddedPrecip,
                               FigS2B_MonthlyGriddedPrecip + theme(legend.position="none"),
                               nrow=1), mylegend, nrow=2,heights=c(10, 1))
```

The variation in annual precipitation increases towards the wetter end of the grid.



# Climate grid
Mean of observation across years for precipitation and tetraterm temperatures (i.e. averages of the months June-September) ± SE (vertical and horizontal error bars) from gridded data for the period of the experiment (5 years, closed symbols) and 30 year normal temperature for the period of 1960-90 (www.met.no, open symbols) confirm the climate grid of the experiment with increasing tetraterm temperature from alpine to boreal sites and increasing precipitation from dry to the wet sites:

```{r, echo=FALSE}

# Filter precipitation and tetraterm temperature
monthlyClimate2 %>% 
  filter(logger == "Temperature" & month(date) %in% 6:9 | logger == "Precipitation") %>% 
  filter(year(date) != 2016) %>% 
  group_by(logger, T_level, P_level, year(date)) %>% 
  summarise(n = n(), mean = mean(value, na.rm = TRUE), sum = sum(value, na.rm = TRUE)) %>% 
  mutate(mean = ifelse(logger == "Precipitation", sum, mean)) %>% 
  group_by(logger, T_level, P_level) %>% 
  summarise(N = n(), Mean = mean(mean), SE = sd(mean, na.rm = TRUE) / sqrt(N)) %>% 
  unite(Mean_SE, Mean, SE, sep = "_") %>% 
  spread(key = logger, value = Mean_SE) %>% 
  separate(col = Precipitation, into = c("P_mean", "P_se"), sep = "_", convert = TRUE) %>% 
  separate(col = Temperature, into = c("T_mean", "T_se"), sep = "_", convert = TRUE) %>%
  select(-N) %>% 
  mutate(logger = "5 year gridded") %>% 
  bind_rows(Longterm) %>%
  ungroup() %>% 
  mutate(P_level = factor(P_level, levels = c(1,2,3,4))) %>% 
  mutate(T_level = factor(T_level, levels = c("Boreal", "Sub-alpine", "Alpine"))) %>% 
  ggplot(aes(x = P_mean, xmin = P_mean - P_se, xmax = P_mean + P_se, y = T_mean, ymin = T_mean -T_se, ymax = T_mean + T_se, color = P_level, shape = T_level, fill = factor(ifelse(logger == "5 year gridded", P_level, logger)))) +
  geom_errorbar() +
  geom_errorbarh() +
  geom_point(aes(size = logger)) +
  labs(x = "Annual precipitation in mm", y = "Tetraterm temperature in °C") +
  scale_color_manual(name = "Precipitation level", values = c("#DFEBF7", "#9DC3E6", "#2E75B6", "#213964")) +
  scale_shape_manual(name = "Temperature level", values = c(25, 21, 24)) +
  scale_fill_manual(name = "Data", values = c("#DFEBF7", "#9DC3E6", "#2E75B6", "white", "#213964")) +
  scale_size_manual(name = "Source", values = c(3, 3.01))+
  guides(fill = "none", size = guide_legend(override.aes = list(shape = c(1, 16)))) +
  theme_minimal(base_size = 10)

```

Colors from light to dark blue are according to increasing precipitation levels 1 = dotted, 2 = dashed, 3 = long-dashed and 4 = solid line. The shape of the symbols represent the temperature levels (boreal: reverse triangle, subalpine: circle, alpine: triangle).
The annual precipitation and tetraterm temperature show that the climate has been both warmer and wetter for the 5 year period of the experiment (closed symbols).



