---
title: "SeedClim-FunCaBClimateData_2008-2017"
author: "Aud Halbritter"
date: "12 2 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE)
```

```{r loadStuff, echo = FALSE}
#Load libraries
library("lubridate")
library("tidyverse")

#load data
load("Temperature.Rdata")
load("Monthly.Temperature_2008-2017.Rdata")
#load("Precipitation.Rdata")
#load("Soilmoisture.Rdata")
```

##  Overview

Climate data from climate logger at 12 the SeedClim / FunCaB sites from the period of 2008 - oct 2017, hourly values


Climate variables are (unit / varible name)

 - Hourly air temperature 2m above the gorund (째C, temp200cm)  
 - Hourly air temperature 30cm above the gorund (째C, temp30cm)  
 - Hourly air temperature 0m above the gorund / in the vegetation (째C, tempabove)  
 - Hourly soil temperature 5cm below the gorund (째C, tempsoil)  
 - Hourly soilmoisture (%, soilmoisture). Always 2 loggers per site! Mean not calculated yet
 - Hourly precipitation (mm, precipitation)  

Daily and monthly mean data for each site is available for all temperature loggers:
- Daily.Temperature_2008-2016.RData
- Monthly.Temperature_2008-2016.RData


## Data processing

The data is provided in txt files.

The raw data is stored in "seedclim climatedata/rawdata by Site" dropbox folder. (will be moved to P drive)

Code for reading the data from the dat files is in 'ReadInSeedClimClimateData.R'. These are archived versions of the code; the working version of the code in the SeedClim-Climate-Data github repository.

Temperature data:
Wrong and problematic data and few data points have been flaged.
Wrong logger names have been corrected.
Soil and aboveground temperature data have been switched if they are wrong.

Wrong variance or mean of data has not been removed yet! Will be done in the near future!

Precipitation and soilmoisture:
Data have not been processed yet!


## Importing the data

The data are provided in .Rdata files. These can be read into R with the `load` command, and ensures that dates etc are correct. The data are in a tidy format: one row - one observation. This code will load the data.
E.g.: load("Temperature.RData", verbose=TRUE)
load("Monthly.Temperature_2008-2016.RData", verbose=TRUE)

```{r loadStuff, eval = FALSE} 
```


```{r eval = FALSE} 
# order sites
temperature <- temperature %>% 
  mutate(site = factor(site, levels = c("Skj", "Gud", "Lav", "Ulv", "Ves", "Ram", "Hog", "Alr", "Ovs", "Arh", "Vik", "Fau")))

#fill missing dates with NA y merging with complete dataset
full_grid <- expand.grid(logger = unique(temperature$logger), site = unique(temperature$site), date = seq(min(temperature$date), max(temperature$date), by = "hour"))

temperature <- left_join(full_grid, temperature) %>% tbl_df()

```


## Plotting the data

Air temperature 2m

```{r}
dd <- temperature %>% 
  filter(logger == "temp200cm")

g <- ggplot(dd, aes(x = date, y = value, color = site)) +
  geom_line() +
  scale_colour_brewer(palette="Paired") +
  facet_wrap(~ site) + 
  theme(legend.position="none") +
  ggtitle("Air temperature 2m")

g
```

Air temperature 30cm

```{r}
dd <- temperature %>% 
  filter(logger == "temp30cm")

g <- ggplot(dd, aes(x = date, y = value, color = site)) +
  geom_line() +
  scale_colour_brewer(palette="Paired") +
  facet_wrap(~ site) + 
  theme(legend.position="none") + ggtitle("Air temperature 30cm")
g
```

Air temperature 0cm

```{r}
dd <- temperature %>% 
  filter(logger == "tempabove")

g <- ggplot(dd, aes(x = date, y = value, color = site)) +
  geom_line() +
  scale_colour_brewer(palette="Paired") +
  facet_wrap(~ site) + 
  theme(legend.position="none") + 
  ggtitle("Air temperature 0cm")
g
```

Soil temperature -5cm

```{r}
dd <- temperature %>% 
  filter(logger == "tempsoil")

g <- ggplot(dd, aes(x = date, y = value, color = site)) +
  geom_line() +
  scale_colour_brewer(palette="Paired") +
  facet_wrap(~ site) + 
  theme(legend.position="none") + 
  ggtitle("Soil temperature -5cm")
g
```


# Monthly temperature

Monthly air temperature 2m
```{r}

monthlyTemperature %>% 
   filter(logger == "temp200cm") %>% 
   ggplot(aes(x = date, y = value, color = site)) +
   geom_line() +
   scale_colour_brewer(palette="Paired") +
   facet_wrap(~ site) + 
   ggtitle("Monthly temperature 2m")

```

Monthly air temperature 30cm
```{r}

monthlyTemperature %>% 
   filter(logger == "temp30cm") %>% 
   ggplot(aes(x = date, y = value, color = site)) +
   geom_line() +
   scale_colour_brewer(palette="Paired") +
   facet_wrap(~ site) + 
   ggtitle("Monthly temperature 30cm")

```

Monthly ground temperature 0cm
```{r}

monthlyTemperature %>% 
   filter(logger == "tempabove") %>% 
   ggplot(aes(x = date, y = value, color = site)) +
   geom_line() +
   scale_colour_brewer(palette="Paired") +
   facet_wrap(~ site) + 
   ggtitle("Monthly temperature 0cm")

```

Monthly soil temperature -5cm
```{r}

monthlyTemperature %>% 
   filter(logger == "tempsoil") %>% 
   ggplot(aes(x = date, y = value, color = site)) +
   geom_line() +
   scale_colour_brewer(palette="Paired") +
   facet_wrap(~ site) + 
   ggtitle("Monthly temperature -5cm")

```




