---
title: "SeedClim-FunCaBClimateData_2008-2018"
author: "Aud Halbritter"
date: "12 2 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE)
```

```{r loadStuff, echo = FALSE}
#Load libraries
library("lubridate")
library("tidyverse")

#load data
load("Temperature.Rdata")
load("Monthly.Temperature_2008-2017.Rdata")
#load("Precipitation.Rdata")
#load("Soilmoisture.Rdata")
```

##  Overview

Climate data from climate logger at 12 SeedClim / FunCaB sites from the period of 2008 - oct 2018, hourly values


Climate variables are (unit / variable name)

 - Hourly air temperature 2m above the ground (째C, temp200cm)  
 - Hourly air temperature 30cm above the ground (째C, temp30cm)  
 - Hourly air temperature 2cm above the ground / in the vegetation (째C, tempabove)  
 - Hourly soil temperature 5cm below the ground (째C, tempsoil)  
 - Hourly soilmoisture (%, soilmoisture). Always 2 loggers per site! Mean not calculated yet
 - Hourly precipitation (mm, precipitation). Data not cleaned, very messy

Daily and monthly mean data for each site is available for all temperature loggers:
- Daily.Temperature_2008-2016.RData
- Monthly.Temperature_2008-2016.RData


## Data processing

#### Data storage and file format
The data is provided in .txt files.
The raw data is stored on the UiB common server (P drive). The data is stored in two different places. We have R code to read in all the data and remove duplicate fiels.

Old repo:
path = "/Volumes/felles/MATNAT/BIO/Felles/007_Funcab_Seedclim/SeedClimClimateData/rawdata by Site/"

New repo:
path = "/Volumes/felles/MATNAT/BIO/Ecological and Environmental Change/SeedClimClimateData/Climate_data_loggers"

ITAS loggers are in separate folders for each year and spring/fall data collection
UTL loggers are in separate folders for each site

Notes on the loggers are stored in .txt files the same folders as the data. File names for notes have to contain "notes", then they are filtered out before reading in the data.

The clean climate data can be found???


#### Code
All code is written in R. These are archived versions of the code; the working version of the code is in the SeedClim-Climate-Data github repository: https://github.com/audhalbritter/SeedClim-Climate-Data

The code to import the climate data from the .txt files is: 'ReadInSeedClimClimateData.R'


#### Data cleaning
Temperature data:
Wrong values are removed (visual estimation of too low or high values)
Problematic data are flagged.
Wrong logger names have been corrected.
Soil and aboveground temperature data have been switched if they are mixed up.

Wrong variance or mean of data has not been removed yet!

Precipitation and soilmoisture:
Data have not been processed yet!


## Importing the data

The data are provided in .Rdata files. These can be read into R with the `load` command, and ensures that dates etc are correct. The data are in a tidy format: one row - one observation. This code will load the data.
E.g.: load("Temperature.RData", verbose=TRUE)
load("Monthly.Temperature_2008-2016.RData", verbose=TRUE)

```{r loadStuff, eval = FALSE} 
```


```{r eval = FALSE} 
# order sites
temperature <- temperature %>% 
  mutate(site = factor(site, levels = c("Skj", "Gud", "Lav", "Ulv", "Ves", "Ram", "Hog", "Alr", "Ovs", "Arh", "Vik", "Fau")))

#fill missing dates with NA y merging with complete dataset
full_grid <- expand.grid(logger = unique(temperature$logger), site = unique(temperature$site), date = seq(min(temperature$date), max(temperature$date), by = "hour"))

temperature <- left_join(full_grid, temperature) %>% tbl_df()

```


## Plotting the data

Air temperature 2m

```{r}
dd <- temperature %>% 
  filter(logger == "temp200cm")

g <- ggplot(dd, aes(x = date, y = value, color = site)) +
  geom_line() +
  scale_colour_brewer(palette="Paired") +
  facet_wrap(~ site) + 
  theme(legend.position="none") +
  ggtitle("Air temperature 2m")

g
```

Air temperature 30cm

```{r}
dd <- temperature %>% 
  filter(logger == "temp30cm")

g <- ggplot(dd, aes(x = date, y = value, color = site)) +
  geom_line() +
  scale_colour_brewer(palette="Paired") +
  facet_wrap(~ site) + 
  theme(legend.position="none") + ggtitle("Air temperature 30cm")
g
```

Air temperature 0cm

```{r}
dd <- temperature %>% 
  filter(logger == "tempabove")

g <- ggplot(dd, aes(x = date, y = value, color = site)) +
  geom_line() +
  scale_colour_brewer(palette="Paired") +
  facet_wrap(~ site) + 
  theme(legend.position="none") + 
  ggtitle("Air temperature 0cm")
g
```

Soil temperature -5cm

```{r}
dd <- temperature %>% 
  filter(logger == "tempsoil")

g <- ggplot(dd, aes(x = date, y = value, color = site)) +
  geom_line() +
  scale_colour_brewer(palette="Paired") +
  facet_wrap(~ site) + 
  theme(legend.position="none") + 
  ggtitle("Soil temperature -5cm")
g
```


# Monthly temperature

Monthly air temperature 2m
```{r}

monthlyTemperature %>% 
   filter(logger == "temp200cm") %>% 
   ggplot(aes(x = date, y = value, color = site)) +
   geom_line() +
   scale_colour_brewer(palette="Paired") +
   facet_wrap(~ site) + 
   ggtitle("Monthly temperature 2m")

```

Monthly air temperature 30cm
```{r}

monthlyTemperature %>% 
   filter(logger == "temp30cm") %>% 
   ggplot(aes(x = date, y = value, color = site)) +
   geom_line() +
   scale_colour_brewer(palette="Paired") +
   facet_wrap(~ site) + 
   ggtitle("Monthly temperature 30cm")

```

Monthly ground temperature 0cm
```{r}

monthlyTemperature %>% 
   filter(logger == "tempabove") %>% 
   ggplot(aes(x = date, y = value, color = site)) +
   geom_line() +
   scale_colour_brewer(palette="Paired") +
   facet_wrap(~ site) + 
   ggtitle("Monthly temperature 0cm")

```

Monthly soil temperature -5cm
```{r}

monthlyTemperature %>% 
   filter(logger == "tempsoil") %>% 
   ggplot(aes(x = date, y = value, color = site)) +
   geom_line() +
   scale_colour_brewer(palette="Paired") +
   facet_wrap(~ site) + 
   ggtitle("Monthly temperature -5cm")

```




