---
title: "GriddedClimateData2009-2017"
author: "Aud Halbritter"
date: "18.01.2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE)
```

```{r loadStuff, echo = FALSE}
#Load libraries
library("ggplot2")
library("dplyr")
library("tidyr")

#load data
load("GriddedDailyClimateData2009-2017.Rdata")
```

##  Overview

Gridded data from met.no (from Ole Einar Tveito, 25.10.2016 and 17.01.2017)
Climate data are for 12 the SeedClim / FunCaB sites from the period of 2009 - 2017, daily values


Climate variables are  
 - Mean daily temperature (°C, Temperature)  
 - Relative air moisture (%, RelAirMoisture)  
 - Mean wind (meter / second, Wind)  
 - Mean cloud cover (in 8 parts, no dimension, CloudCover)  
 - Precipitation (mm, Precipitation)

Monthly and annual mean data for each site is available:
GriddedMonthly_AnnualClimate2009-2017.RData


## Data processing

The data was provided in dat files.

The raw data is stored on P:/FELLES/MATNAT/BIO/Ecological and Environmental Change/SeedClimClimateData/met-data/GriddedClimateData2009-2017

Code for reading the data from the dat files is in 'ReadInGriddedData.R'. These are archived versions of the code; the working version of the code in the SeedClim-Climate-Data github repository.

Site names were changed from a ID that was used to extranct the data to the 3 letter site names used in SeedClim (Gud, Lav, Vik,...). The Code for the ID's can be found on P drive.


## Importing the data

The data are provided in .Rdata files. These can be read into R with the `load` command, and ensures that dates etc are correct. The data are in a tidy format: one row - one observation. This code will load the data.
E.g.: load("GriddedDailyClimateData2009-2017.RData", verbose=TRUE)
load("GriddedMonthly_AnnualClimate2009-2017.RData", verbose=TRUE)

```{r loadStuff, eval = FALSE} 
```


## Plotting the data

Temperature

```{r}
ggplot(climate, aes(x = Date, y = Temperature, color = Site)) +
  geom_line() +
  scale_colour_brewer(palette="Paired") +
  facet_wrap(~ Site) + 
  theme(legend.position="none") +
  ggtitle("Temperature")
```



RelAirMoisture

```{r}
ggplot(climate, aes(x = Date, y = RelAirMoisture, color = Site)) +
  geom_line() +
  scale_colour_brewer(palette="Paired") +
  facet_wrap(~ Site) + 
  theme(legend.position="none") +
  ggtitle("Relative Air Moisture")
```


Wind
```{r}
ggplot(climate, aes(x = Date, y = Wind, color = Site)) +
  geom_line() +
  scale_colour_brewer(palette="Paired") +
  facet_wrap(~ Site) + 
  theme(legend.position="none") +
  ggtitle("Wind")

```


CloudCover
```{r}
ggplot(climate, aes(x = Date, y = CloudCover, color = Site)) +
  geom_line() +
  scale_colour_brewer(palette="Paired") +
  facet_wrap(~ Site) + 
  theme(legend.position="none") +
  ggtitle("Cloud Cover")
```


Precipitation
```{r}
ggplot(climate, aes(x = Date, y = Precipitation, color = Site)) +
  geom_line() +
  scale_colour_brewer(palette="Paired") +
  facet_wrap(~ Site) + 
  theme(legend.position="none") +
  ggtitle("Precipitation")

```



