---
title: "GriddedClimateData2009-2019"
author: "Aud Halbritter"
date: "19.12.2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, cache = TRUE)
```

```{r loadStuff, echo = FALSE}
#Load libraries
library("tidyverse")
library("lubridate")

# read in data to R
climate <- read_csv(file = "GriddedDailyClimateData2009-2019.csv", col_names = TRUE)

```

##  Overview

Gridded data from met.no (from Ole Einar Tveito, 25.10.2016, 17.01.2017, 16.12.2019)
Climate data are for 12 the SeedClim / FunCaB sites from the period of 2009 - 2019, daily values


Climate variables are  
 - Mean daily temperature (°C, Temperature)  
 - Relative air moisture (%, RelAirMoisture)  
 - Mean wind (meter / second, Wind)  
 - Mean cloud cover (in 8 parts, no dimension, CloudCover)  
 - Precipitation (mm, Precipitation)

Code for calculating monthly and annual mean data for each site is described below.


## Data processing

The data was provided in dat files.

The raw data is stored on P:/FELLES/MATNAT/BIO/Ecological and Environmental Change/SeedClimClimateData/met-data/GriddedClimateData2009-2017

Code for reading the data from the dat files is in 'ReadInGriddedData.R'. These are archived versions of the code; the working version of the code in the SeedClim-Climate-Data github repository.

Site names were changed from a ID that was used to extranct the data to the 3 letter site names used in SeedClim (Gud, Lav, Vik,...). The Code for the ID's can be found on P drive.


## Importing the data

The data are provided as a csv file on OSF (https://osf.io/npfa9/). These can be downloaded to your R project using the get_file function from the downloader package. The first time osfr and downloader need to be installed (see below). Downloader will download the data from OSF and create a folder called data_cleaned.
The data are in a tidy format: one row - one observation. This code will load the data.

```{r importData, eval = FALSE} 

#remotes::install_github("centerforopenscience/osfr")
#devtools::install_github("Between-the-Fjords/downloader")
library("osfr")
library("downloader")

# Download data from OSF
get_file(node = "npfa9", 
         file = "GriddedDailyClimateData2009-2019.csv", 
         path = "data_cleaned")

# read in data to R
climate <- read_csv(file = "data_cleaned/GriddedDailyClimateData2009-2019.csv", col_names = TRUE)

```


## Calculate monthly and annual values

Code to calculate monthly and annual climate data.

```{r}
# Calculate Monthly Mean
monthlyClimate <- climate %>%
  select(-Year, -Month, -Day) %>% 
  gather(key = Logger, value = value, -Site, -Date) %>% 
  mutate(dateMonth = dmy(paste0("15-",format(Date, "%b.%Y")))) %>%
  group_by(dateMonth, Logger, Site) %>%
  summarise(n = n(), value = mean(value), sum = sum(value)) %>% 
  mutate(value = ifelse(Logger == "Precipitation", sum, value)) %>% 
  select(-n, -sum)


# Calculate Annual Means
annualClimate <- monthlyClimate %>% 
  mutate(dateYear = year(dateMonth)) %>%
  group_by(dateYear, Logger, Site) %>%
  summarise(annualMean = mean(value)) %>% 
  spread(key = Logger, value = annualMean)

```


## Plotting the data

Temperature

```{r}
p <- ggplot(climate, aes(x = Date, y = Temperature, color = Site)) +
  geom_line() +
  scale_colour_brewer(palette="Paired") +
  facet_wrap(~ Site) + 
  theme(legend.position="none") +
  ggtitle("Temperature")
p
```


RelAirMoisture

```{r}
p + aes(y = RelAirMoisture) +
  ggtitle("Relative Air Moisture")
```


Wind
```{r}
p + aes(y = Wind) +
  ggtitle("Wind")

```


CloudCover
```{r}
p + aes(y = CloudCover) +
  ggtitle("Cloud Cover")
```


Precipitation
```{r}
p + aes(y = Precipitation) +
  ggtitle("Precipitation")

```



