---
title: "SeedClim Climate"
author: "Richard J. Telford"
date: '2016-04-29'
output: ioslides_presentation
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(fig.path='Figs/', echo=FALSE, warning=FALSE, message=FALSE)
```

```{r importChunk, results = "hide", fig.keep='none', cache = TRUE}
source("klima.R")
```

```{r packages}
library(ggplot2)
library(plyr)
library(reshape2)
library(lubridate)
```
## Data

Loggers

 - temperature 2m
 - temperature 30cm
 - temperature surface
 - temperature soil
 - soil moisture
 - precipitation

Gridded data from met.no

 - 48 GB of temperature data
 - precipitation data 


## Precipitation data needs careful cleaning

```{r precipitationChunk}
ggplot(clim[[12]], aes(x = date, y = precip)) + geom_path() + ggtitle("Vikesland")
```

## Alrust {.smaller}
```{r longSeriesChunk}
  clim1a <- melt(clim[[1]], id.vars = c("Site", "date"), measure.vars = c("temp2m", "temp30cm", "tempGroundGrass", "tempSoil"))
  g <- ggplot(clim1a, aes(x = date, y = value, colour = variable)) + geom_line() + labs(y = "Temperature °C")
print(g)

```
<div class="columns-2">
 - Missing data
 - Unearthed Soil temperature loggers
 - Occasional spurious readings
</div>


## Daily cycles

```{r dailyCyclesChunk}
g <- ggplot(subset(clim1, year(date) == 2013 & month(date) == 7), aes(x = hour(date) + minute(date)/60, y = value, colour = variable, group = factor(day(date)))) + geom_path() + labs(x = "Time, hours", y = "Temperature °C") + facet_wrap(~variable)
print(g)
```

## Daily Differences from 2m temperature

Alrust July 2013
```{r dailyDeltaChunk}
g <- ggplot(subset(clim1d, year(date) == 2013 & month(date) == 7 & variable != "temp2m"), aes(x = hour(date) + minute(date)/60, y = value, colour = variable, group = factor(day(date)))) + geom_path() + labs(x = "Time, hours", y = "Temperature °C")+ facet_wrap(~variable)
print(g)
```



## Across the grid

July Temperatures
```{r, echo = FALSE, message = FALSE, warning=FALSE }
g <- ggplot(subset(climMeans, format(date, "%b") == "Jul"), aes(x = year(date), y = temp2m, colour = site)) + geom_line() + sc + labs(x = "Year", y = "Temperature °C")
print(g)
```

## Across the grid

August Temperatures
```{r, echo = FALSE, message = FALSE, warning=FALSE }
g <- ggplot(subset(climMeans, format(date, "%b") == "Aug"), aes(x = year(date), y = temp2m, colour = site)) + geom_line() + sc + labs(x = "Year", y = "Temperature °C")
print(g)
```

## Month mean differences 2m - 30cm

```{r, echo = FALSE, message = FALSE, warning=FALSE }
ggplot(climMeans, aes(x = temp2m, y = temp30cm, colour = site)) + geom_point() + geom_abline(slope = 1, intercept = 0) + sc + labs(x = "2 m Temperature °C", y = "30 cm Temperature °C")
```
<div class="columns-2">
- snow
 - radiation frost
 - data issues
</div>


## delta 30-2m for all sites

```{r, echo = FALSE, warning=FALSE}
clim1 <- ldply(clim, function(x){
    x$delta30 <- x$temp30cm - x$temp2m
    x
})
  
  clim1m <- melt(clim1[,c(2:4, 19)], id = c("Site", "date"))
  g <-ggplot(subset(clim1m,  month(date) == 7 & variable == "delta30"), aes(x = hour(date) + minute(date)/60, y = as.numeric(value), colour = factor(year(date)), group = format(date, "%d-%m-%Y"))) + geom_path() + labs(x = "Time, hours", y = "Temperature °C") + facet_wrap(~Site) + scale_colour_discrete(name = "Year")
print(g)
```
